State of the Practice for Lattice Boltzmann Method Software

### 2022-03-04 StateOfPractice/Papers/LatticeBoltz_SOP-arXiv/LatticeBoltz_SOP.tex
Edits to abstract and conclusions

    We analyze the state of software development practice for Lattice Boltzmann
    solvers and find by quantitatively and qualitatively measuring and comparing
    24 software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP). Our ranking was roughly
    consistent with ranking the repositories by the number of GitHub stars,
    although the number one project by star count (Sailfish) is ranked 9th by
    our methodology. Our finding is that the state of the practice is healthy
    with 67\% of the measured packages ranking in the top 5 for at least one
    quality, the majority of LBM generated artifacts corresponding to general
    recommendations from research software developers, common utilization (67\%
    of packages) of version control and adoption of a quasi-agile development
    process.Areas of best practice to potentially improve include adoption of
    continuous integration, API documentation and enforcement of programming
    style guides.  We interviewed 4 developers to gain insight into their
    current pain points. Identified challenges include lack of development time,
    lack of funding, and difficulty with ensuring correctness.  Developers are
    addressing these pain points by designing for change, circumventing the
    oracle problem and prioritizing documentation and usability.  For future the
    best practices of the future, we suggest the community consider: employing
    linters, conducing rigorous peer reviews, writing and submitting more papers
    on software, growing the number of contributors and augmenting the theory
    manuals to include more requirements relevant information.

@@

To help users select the right LBM package for their needs, and to potentially
provide direction for future LBM software development, our goal was to analyze
the current state of LBM software development.  We aimed to determine how well
best practices are followed and to understand current developer pain points.  We
compared the state of the practice for LBM software to the development practices
recommended and employed for research software in general. To improve the
quality of LBM software we looked for examples where developers have been
successful so that we can share these with the LBM community, and with the
broader research software community.  We also aimed to identify potential
approaches for mitigating the identified pain points in the future.

@@

To advance the state of the practice beyond current best practices toward future
best practices, we interviewed 4 developers to learn about their pain points
(Section~\ref{painpoints}).  The identified pain points include lack of
development time, lack of software development experience, lack of incentive and
funding, lack of external support, technology hurdles, ensuring correctness,
usability, technical debt and documentation quality.  We found
(Section~\ref{Sec_AddressConcerns} that developers are currently creating
practices to address these pain points, including: designing for change,
circumventing the oracle testing problem, and prioritizing documentation and
usability.  To further plan for future best practices, we identified eight
potential directions (Section~\ref{Sec_Recommendations}), in order of increasing
novelty:

@@

Future work can be done to address the threats to validity mentioned in
Section~\ref{threats}.  For instance, the measures listed in our measurement
template could be broadened to better capture some qualities. For example,
usability experiments and performance benchmarks could be incorporated into the
assessment.  Section~\ref{threats} also mentions that some pertinent information
on continuous integration and unit testing was not specified in public
artifacts. Adding further questions to the interview guide regarding the
measures that are on the measurement template could reduce instances of
incomplete data. This additional interview data could be analyzed and
incorporated into the AHP ranking, ensuring quality designations more accurate
represent the true quality of the software packages. In the future we could also
expand the interview questions to go beyond reproducibility to also discuss
replicability. Furthermore, in future interviews we would like to clarify the
distinction between usability and understandability.

Our domain analysis (Section~\ref{SecDomainAnalysis} was relatively shallow.  In
the future we would like to dig more deeply into this analysis.  We would like
to produce tables of commonalities, variabilities, and parameters of variation
for the family of LBM solvers. We could follow the template from
\citep{smith2008commonality}, or adopt the analysis techniques from
\citep{weiss1998commonality}. \cite{SmithAndChen2004} and
\cite{SmithMcCutchanAndCarette2017} are examples of a commonality analysis for a
family of mesh generating software and a family of material models,
respectively.

>>>>

    We analyze the state of software development practice for Lattice Boltzmann
    solvers by quantitatively and qualitatively measuring and comparing 24
    software packages for 10 software qualities (installability, correctness/
    verifiability, reliability, robustness, usability, maintainability,
    reusability, understandability, visibility/transparency and
    reproducibility). Our reproducible analysis method employs a measurement
    template (containing 108 measures that are manually and automatically
    extracted from the project repository) and developer interviews (consisting
    of a set of 20 questions). From the measurement results, we ranked the
    software using the Analytic Hierarchy Process (AHP). Our ranking was roughly
    consistent with ranking the repositories by the number of GitHub stars,
    although the number one project by star count (Sailfish) is ranked 9th by
    our methodology. Our finding is that the state of the practice is healthy
    with 67\% of the measured packages ranking in the top 5 for at least one
    quality, the majority of LBM generated artifacts corresponding to general
    recommendations from research software developers, common utilization of
    version control (67\% of packages) and adoption of a quasi-agile development
    process.  Areas of best practice to potentially improve include adoption of
    continuous integration, API documentation and enforcement of programming
    style guides.  We interviewed 4 developers to gain insight into their
    current pain points. Identified challenges include lack of development time,
    lack of funding, and difficulty with ensuring correctness.  Developers are
    addressing these pain points by designing for change, circumventing the
    oracle problem and prioritizing documentation and usability.  For future the
    best practices, we suggest the community consider the following: employing
    linters, conducing rigorous peer reviews, writing and submitting more papers
    on software, growing the number of contributors and augmenting the theory
    manuals to include more requirements relevant information.

@@

To help users select the right LBM package for their needs, and to potentially
provide direction for future LBM software development, our goal was to analyze
the current state of LBM software development.  We determined how well best
practices are followed and we investigated current developer pain points.  We
compared the state of the practice for LBM software to the development practices
recommended and employed for research software in general. To improve the
quality of LBM software we looked for examples where developers have been
successful so that we can share these with the LBM community, and with the
broader research software community.  We also aimed to identify potential
approaches for mitigating the identified pain points in the future.

@@

To advance the state of the practice beyond current best practices toward future
best practices, we interviewed 4 developers to learn about their pain points
(Section~\ref{painpoints}).  The identified pain points include lack of
development time, lack of software development experience, lack of incentive and
funding, lack of external support, technology hurdles, ensuring correctness,
usability, technical debt and documentation quality.  We found
(Section~\ref{Sec_AddressConcerns} that developers are currently creating
practices to address these pain points, including: designing for change,
circumventing the oracle testing problem, and prioritizing documentation and
usability.  To plan for future best practices, we identified eight
potential directions (Section~\ref{Sec_Recommendations}), in order of increasing
novelty:

@@

Future work can be done to address the threats to validity mentioned in
Section~\ref{threats}.  For instance, the measures listed in our measurement
template could be broadened to better capture some qualities. For example,
usability experiments and performance benchmarks could be incorporated into the
assessment.  Section~\ref{threats} also mentions that some pertinent information
on continuous integration and unit testing was not specified in public
artifacts. Adding further questions to the interview guide regarding the
measures that are on the measurement template could reduce instances of
incomplete data. This additional interview data could be analyzed and
incorporated into the AHP ranking, ensuring quality designations more accurately
represent the true quality of the software packages. In the future we could also
expand the interview questions to go beyond reproducibility to also discuss
replicability. Furthermore, in future interviews we would look to clarify the
distinction between usability and understandability
(Section~\ref{softwarequalities}), since the distinction confused some
participants.

Our domain analysis (Section~\ref{SecDomainAnalysis} was relatively shallow.  In
the future we would like to dig more deeply into this analysis.  We would like
to produce tables of commonalities, variabilities, and parameters of variation
for the family of LBM solvers. We could follow the template from
\citep{smith2008commonality}, or adopt the analysis techniques from
\citep{weiss1998commonality}. \cite{SmithAndChen2004} and
\cite{SmithMcCutchanAndCarette2017} provide examples that could be followed for
a commonality analysis of a family of mesh generating software and a family of
material models, respectively.
